sdmTMB: an R package for fast, flexible, and user-friendly generalized linear mixed effects models with spatial and spatiotemporal random fields

### 2023-09-08 doc/paper-jss/paper-jss.Rmd
Comment out old examples

```{r sdmTMB-install, eval=FALSE, echo=TRUE}
install.packages("sdmTMB", dependencies = TRUE)
```

@@

```{r sdmTMB-lib0, eval=FALSE, echo=FALSE}
# library("sdmTMB")
# # simplify for paper:
pcod_original <- pcod # save it for later
# pcod <- dplyr::select(pcod, year, density, depth, lon, lat) |>
#   tibble::as_tibble()

```

```{r sdmTMB-lib, eval=FALSE, echo=TRUE}
library("sdmTMB")
dplyr::select(pcod, year, density, depth, lon, lat) |>
  head(n = 3)
```

@@

```{r pcod-utms-eval, echo=TRUE, eval=FALSE}
pcod <- add_utm_columns(pcod, c("lon", "lat"), units = "km")
```

@@

```{r pcod-head2, echo=TRUE, eval=FALSE}
head(pcod, n = 3)
```

@@

```{r, echo=TRUE, eval=FALSE, cache=TRUE}
mesh <- make_mesh(pcod, xy_cols = c("X", "Y"), cutoff = 10)
```

@@

```{r pcod-eg1-fit, warning=FALSE, message=FALSE, echo=TRUE, eval=FALSE, cache=TRUE}
fit <- sdmTMB(
  density ~ s(depth),
  data = pcod,
  family = tweedie(link = "log"),
  mesh = mesh,
  spatial = "on"
)
```

@@

```{r pcod-eg1-summary,eval=FALSE, echo=TRUE}
fit
```

@@

```{r pcod-eg1-tidy-fe, eval=FALSE, echo=TRUE}
tidy(fit, conf.int = TRUE)
```

```{r pcod-eg1-tidy-re, eval=FALSE, echo=TRUE}
tidy(fit, effects = "ran_pars", conf.int = TRUE)
```

@@

```{r pcod-eg1-sanity, eval=FALSE, echo=TRUE}
sanity(fit)
```

@@

```{r pcod-eg1-suppress-depth2, eval=FALSE, echo=FALSE}
qcs_grid_original <- qcs_grid
qcs_grid$depth_scaled <- NULL # for simplicity in paper
qcs_grid$depth_scaled2 <- NULL # for simplicity in paper
```

```{r qcs-head, echo=FALSE}
head(qcs_grid, n = 3)
```

```{r pcod-eg1-predict, echo=FALSE, eval=TRUE, cache=TRUE}
p <- predict(fit, newdata = qcs_grid)
```

@@

```{r pcod-p-tibble, eval=FALSE}
p <- tibble::as_tibble(p)
```

```{r pcod-head-predict, echo=TRUE, eval=FALSE}
head(p, n = 3)
```

@@

```{r time1, cache=TRUE, eval=FALSE}
t1 <- Sys.time()
```

```{r fit-tv, eval=FALSE, echo=TRUE, cache=TRUE, results='hide'}
fit_spatiotemporal <- sdmTMB(
  density ~ 0 + s(depth, k = 5),
  time_varying = ~ 1,
  time_varying_type = "rw",
  family = tweedie(link = "log"),
  data = pcod,
  mesh = mesh,
  time = "year",
  spatial = "on",
  spatiotemporal = "iid",
  anisotropy = TRUE,
  silent = FALSE
)
```

```{r time2, cache=TRUE, eval=FALSE}
t2 <- Sys.time()
```

```{r fit-tv-print, eval=FALSE, echo=TRUE}
fit_spatiotemporal
```

@@

```{r pcod-restore, echo=FALSE, eval=FALSE}
pcod <- pcod_original
qcs_grid <- qcs_grid_original
```

@@

```{r owl-fit-old, warning=FALSE, message=FALSE, cache=TRUE, echo=TRUE, eval=FALSE}
mesh <- make_mesh(snow, xy_cols = c("X", "Y"), cutoff = 1.5)
fit_owls <- sdmTMB(
  count ~ nao + (1 | year_factor),
  spatial_varying = ~ nao,
  family = nbinom2(link = "log"),
  data = snow,
  mesh = mesh,
  time = "year",
  spatial = "on",
  spatiotemporal = "iid"
)
```

>>>>

```{r sdmTMB-install, eval=FALSE, echo=TRUE}
# install.packages("sdmTMB", dependencies = TRUE)
```

@@

```{r sdmTMB-lib0, eval=FALSE, echo=FALSE}
# # library("sdmTMB")
# # # simplify for paper:
# pcod_original <- pcod # save it for later
# # pcod <- dplyr::select(pcod, year, density, depth, lon, lat) |>
# #   tibble::as_tibble()
```

```{r sdmTMB-lib, eval=FALSE, echo=TRUE}
# library("sdmTMB")
# dplyr::select(pcod, year, density, depth, lon, lat) |>
#   head(n = 3)
```

@@

```{r pcod-utms-eval, echo=TRUE, eval=FALSE}
# pcod <- add_utm_columns(pcod, c("lon", "lat"), units = "km")
```

@@

```{r pcod-head2, echo=TRUE, eval=FALSE}
# head(pcod, n = 3)
```

@@

```{r, echo=TRUE, eval=FALSE, cache=TRUE}
# mesh <- make_mesh(pcod, xy_cols = c("X", "Y"), cutoff = 10)
```

@@

```{r pcod-eg1-fit, warning=FALSE, message=FALSE, echo=TRUE, eval=FALSE, cache=TRUE}
# fit <- sdmTMB(
#   density ~ s(depth),
#   data = pcod,
#   family = tweedie(link = "log"),
#   mesh = mesh,
#   spatial = "on"
# )
```

@@

```{r pcod-eg1-summary,eval=FALSE, echo=TRUE}
# fit
```

@@

```{r pcod-eg1-tidy-fe, eval=FALSE, echo=TRUE}
# tidy(fit, conf.int = TRUE)
```

```{r pcod-eg1-tidy-re, eval=FALSE, echo=TRUE}
# tidy(fit, effects = "ran_pars", conf.int = TRUE)
```

@@

```{r pcod-eg1-sanity, eval=FALSE, echo=TRUE}
# sanity(fit)
```

@@

```{r pcod-eg1-suppress-depth2, eval=FALSE, echo=FALSE}
# qcs_grid_original <- qcs_grid
# qcs_grid$depth_scaled <- NULL # for simplicity in paper
# qcs_grid$depth_scaled2 <- NULL # for simplicity in paper
```

```{r qcs-head, echo=FALSE}
# head(qcs_grid, n = 3)
```

```{r pcod-eg1-predict, echo=FALSE, eval=TRUE, cache=TRUE}
# p <- predict(fit, newdata = qcs_grid)
```

@@

```{r pcod-p-tibble, eval=FALSE}
# p <- tibble::as_tibble(p)
```

```{r pcod-head-predict, echo=TRUE, eval=FALSE}
# head(p, n = 3)
```

@@

```{r time1, cache=TRUE, eval=FALSE}
# t1 <- Sys.time()
```

```{r fit-tv, eval=FALSE, echo=TRUE, cache=TRUE, results='hide'}
# fit_spatiotemporal <- sdmTMB(
#   density ~ 0 + s(depth, k = 5),
#   time_varying = ~ 1,
#   time_varying_type = "rw",
#   family = tweedie(link = "log"),
#   data = pcod,
#   mesh = mesh,
#   time = "year",
#   spatial = "on",
#   spatiotemporal = "iid",
#   anisotropy = TRUE,
#   silent = FALSE
# )
```

```{r time2, cache=TRUE, eval=FALSE}
# t2 <- Sys.time()
```

```{r fit-tv-print, eval=FALSE, echo=TRUE}
# fit_spatiotemporal
```

@@

```{r pcod-restore, echo=FALSE, eval=FALSE}
# pcod <- pcod_original
# qcs_grid <- qcs_grid_original
```

@@

```{r owl-fit-old, warning=FALSE, message=FALSE, cache=TRUE, echo=TRUE, eval=FALSE}
# mesh <- make_mesh(snow, xy_cols = c("X", "Y"), cutoff = 1.5)
# fit_owls <- sdmTMB(
#   count ~ nao + (1 | year_factor),
#   spatial_varying = ~ nao,
#   family = nbinom2(link = "log"),
#   data = snow,
#   mesh = mesh,
#   time = "year",
#   spatial = "on",
#   spatiotemporal = "iid"
# )
```
